Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-19 ms.tex
Update on Overleaf.

\begin{abstract}
%{\bf Ignore: Ionizing radiation from recently formed massive stars is important in destroying cold, dense gas around new regions of star formation that would otherwise continue to form stars. Applying a local prescription for this pre-supernova feedback can be sufficient in matching star formation rates in a simulation with full radiative transfer on short timescales. However, the long-range effects of ionizing radiation are important in producing a volume-filling diffuse, warm-ionized medium in the ISM and in the disk-halo interface. This is essential for allowing supernovae to drive significant gas outflows that travel throughout and beyond the virial radius of the galaxy. A simulation with local radiation feedback only is unable to drive significant outflows as the supernovae are trapped well within the inner halo by higher density cold and warm neutral gas.}
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby simulations including self-gravity and radiative cooling alone produce galaxies with far too many stars. This problem has been addressed by employing various models of feedback physics which produces self-regulating star formation in galaxies. Recently, it has become clear that simple models for stellar feedback, which account for supernovae (SNe) alone, are insufficient to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Additional processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging.

@@

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Stromgren approximation whereby the particles / cells within the Stromgren radius a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2016}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds. Although the approximate radiation feedback models are capable of driving significant galactic outflows from galaxies, it is unclear if they capture the full effects of stellar ionizing radiation.

In this work we present a comparison of three simulations of the evolution of an isolated, low mass dwarf galaxy that account for detailed stellar feedback. Our fiducial model uses an adaptive ray-tracing radiative transfer method to follow stellar ionizing radiation. We test its effectiveness in regulating star formation and driving galactic winds by comparison to a simulation without any ionizing radiation and a simulation with only localized radiation around each star. We discuss our methods in Section~\ref{sec:methods} and present our results in Section~\ref{sec:results}.

@@

We include a UV background, tabulated metal line cooling, and a 9 species non-equilibrium chemistry solver using \textsc{Grackle} \citep{GrackleMethod}. Star formation occurs stochastically in cold, dense regions ($n > 200$~cm$^{-3}$, $T < 200$~K) by sampling the \cite{Salpeter1955} IMF from 1~M$_{\odot}$ to 100~M$_{\odot}$ and depositing \textit{individual} star particles over this mass range. For each star, we include feedback from stellar winds, AGB winds, FUV and LW band radiation which drives photoelectric heating and H$_2$ dissociation respectively, HI and HeI ionizing radiation, and core collapse and Type Ia SNe using thermal energy injection. FUV and LW band radiation are both taken to be optically thin, with local (cell-by-cell) attenuation alone. Ionizing radiation is followed using radiative transfer, as discussed below. \textit{Maybe can leave this out:} Stellar lifetimes, surface gravity, effective temperature, and radii are set by the initial stellar mass and metallicity through interpolation over the PARSEC \citep{Bressan2012} zero age main sequence values. These values for surface gravity, effective temperature, and radii are used to set the FUV, LW, and ionizing photon fluxes from each star through interpolation on the OSTAR2002 \citep{LanHubeny2003} grid. We only follow the radiation from stars above 8~M$_{\odot}$. 
%tellar wind and supernovae mass loss are determined with the  NuGrid yield tables for $M_* < 25$~M$_{\odot}$ and from Slemer et. al. (in prep) for $M_* > 25$~M$_{\odot}$.

@@

This can be confirmed by examining the mass loading factor and metal retention fraction (the fraction of produced metals retained within the disk of the galaxy) in each simulation. As shown in Figure~\ref{fig:metal_retention}, the mass loading factor in the fiducial run is a factor of XXX higher at 0.25~R$_{\rm vir}$ than the shortrad simulation, and is the only simulation with any significant outflow beyond the virial radius. This is especially interesting considering the factor 5 increase in SFR of the noRT run. The noRT run does have a significant amount of gas mass outflowing at XXXX, but note that its low mass loading factor is due primarily to the significantly larger SFR in this galaxy. Radiation feedback, both locally and over longer ranges, is important in determining the ability for SN to drive significant outflows. We discuss the reasons for this in the following section.

\begin{figure}
\centering
\includegraphics[width=0.98\linewidth]{metal_retention}
\caption{The fraction of metals contained within the ISM of each galaxy (blue) and the fraction ejected beyond the virial radius (orange). As in Figure~\ref{fig:sfr_mass_evolution}, the fiducial run is given as solid lines, while the noRT and shortrad are dash-dotted and dashed respectively.}
\label{fig:metal_retention}
\end{figure}

\textbf{We have room for one more figure and a little more text to explain it. Possibly a comparison of mass / volume fractions in certain phases? But maybe not necessary....}

\section{Discussion} \label{sec:discussion}
Again, it is known that accounting for feedback from stellar radiation plays a significant role in determining the ability for SN energy to couple to the ISM and therefore drive outflows. However, the specific importance of localized ionization vs. ionization from a diffuse radiation field far from a single star has not been examined. As shown above, accounting for stellar radiation feedback locally alone is insufficient to describe the long-term evolution of an isolated dwarf galaxy. To explore the cause of this difference we present two panel-plots in Figure~\ref{fig:panel1} and Figure~\ref{fig:panel2} which compare the gas number density (left), temperature (middle), and hydrogen ionization fraction (right) in edge-on slices in each of our simulations at two different times.\footnote:{See ({\it insert link here once posted}) for a movie of this comparison.}

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{\textbf{In the final version of this figure I will add labels for each row marking the simulation, and a line giving the scale of each figure.} Edge-on slices of each simulation showing gas number density (left), temperature (middle), and hydrogen ionization fraction (right) 17~Myr after the formation of the first star in each run.}
\label{fig:panel1}
\end{figure*}

@@

\section{Conclusion}  \label{sec:conclusion}
In agreement with previous works, we find that stellar radiation is an important source of feedback that helps regulate star formation in dwarf galaxies. (more)

>>>>

\begin{abstract}
\textit{Will complete at end}
%{\bf Ignore: Ionizing radiation from recently formed massive stars is important in destroying cold, dense gas around new regions of star formation that would otherwise continue to form stars. Applying a local prescription for this pre-supernova feedback can be sufficient in matching star formation rates in a simulation with full radiative transfer on short timescales. However, the long-range effects of ionizing radiation are important in producing a volume-filling diffuse, warm-ionized medium in the ISM and in the disk-halo interface. This is essential for allowing supernovae to drive significant gas outflows that travel throughout and beyond the virial radius of the galaxy. A simulation with local radiation feedback only is unable to drive significant outflows as the supernovae are trapped well within the inner halo by higher density cold and warm neutral gas.}
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby the action of self-gravity and radiative cooling alone produces galaxies with far too many stars. This problem has been addressed by employing various models of stellar feedback physics which are capable of generating self-regulating star formation in galaxies. Recently, it has become clear that the simplest models for stellar feedback, which account for supernovae (SNe) alone, are insufficient \citep{Smith2018} to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging. As a result, the relative importance of these additional physics is still uncertain.

@@

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Str{\"o}mgren approximation whereby the particles / cells within the Str{\"o}mgren radius around a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2016}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. Although the approximate radiation feedback models are capable of driving significant galactic outflows, it is unclear if they capture the full effects of stellar ionizing radiation. 
%The 
%It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds.

In this work we present a comparison of three simulations of the evolution of an isolated, low mass dwarf galaxy that account for detailed stellar feedback. Our fiducial model uses an adaptive ray-tracing radiative transfer method to follow stellar ionizing radiation. We test its effectiveness in regulating star formation and driving galactic winds by comparison to a simulation without ionizing radiation and a simulation with only localized radiation around each star. We discuss our methods in Section~\ref{sec:methods} and present our results in Section~\ref{sec:results}.

@@

We include a UV background, tabulated metal line cooling, and a 9 species non-equilibrium chemistry solver using \textsc{Grackle} \citep{GrackleMethod}. Star formation occurs stochastically in cold, dense regions ($n > 200$~cm$^{-3}$, $T < 200$~K) by sampling the \cite{Salpeter1955} IMF from 1~M$_{\odot}$ to 100~M$_{\odot}$ and depositing \textit{individual} star particles over this mass range. For each star, we include feedback from stellar winds, AGB winds, FUV and LW band radiation which drives photoelectric heating and H$_2$ dissociation respectively, HI and HeI ionizing radiation, and core collapse and Type Ia SNe using thermal energy injection. FUV and LW band radiation are both taken to be optically thin, with local (cell-by-cell) attenuation alone. Ionizing radiation is followed using radiative transfer, as discussed below. Stellar lifetimes, surface gravity, effective temperature, and radii are set by the initial stellar mass and metallicity through interpolation over the PARSEC \citep{Bressan2012} zero age main sequence values. These values for surface gravity, effective temperature, and radii are used to set the FUV, LW, and ionizing photon fluxes from each star through interpolation on the OSTAR2002 \citep{LanzHubeny2003} grid. We only follow the radiation from stars with masses above 8~M$_{\odot}$. 
%tellar wind and supernovae mass loss are determined with the  NuGrid yield tables for $M_* < 25$~M$_{\odot}$ and from Slemer et. al. (in prep) for $M_* > 25$~M$_{\odot}$.

@@

This can be confirmed by examining the metal retention fraction (the fraction of produced metals retained within the disk of the galaxy) and mass outflow rates in each simulation. As shown in Figure~\ref{fig:metal_retention}, the mass outflow rate in the fiducial run peaks at an order of magnitude higher at 0.25~R$_{\rm vir}$ than the shortrad simulation, declining only due to a comparative drop off in star formation. Given this, the fiducial simulation has an even higher mass loading factor, since the shortrad simulation has a typically higher SFR. The fiducial run is also the only simulation with any significant outflow beyond the virial radius, though we note the noRT run has not been simulated long enough for gas to have reached the virial radius. The outflow in noRT is nearly the same as the fiducial run, but it requires a 5 times higher supernova rate in this simulation to match the same outflow seen with full stellar radiation feedback. Radiation feedback, therefore allows SN to be substantially more effective in driving outflows. We provide a physical intuition for these differences in the following section.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{metal_retention}\\
\includegraphics[width=0.95\linewidth]{mass_loading_comparison}
\caption{{\bf Top:} The fraction of metals contained within the ISM of each galaxy (blue) and the fraction ejected beyond the virial radius (orange). As in Figure~\ref{fig:sfr_mass_evolution}, the fiducial run is given as solid lines, while the noRT and shortrad are dash-dotted and dashed respectively. {\bf Bottom:} The mass outflow rate for each run at 0.25~R$_{\rm vir}$ (solid) and R$_{\rm vir}$ (dashed).}
\label{fig:metal_retention}
\end{figure}

\textbf{We may have room for one more figure and a little more text to explain it. Possibly a comparison of mass / volume fractions in certain phases? But maybe not necessary....}

\section{Discussion} \label{sec:discussion}
Again, it is known that accounting for feedback from stellar radiation plays a significant role in determining the ability for SN energy to couple to the ISM and therefore drive outflows. However, the specific importance of localized ionization vs. ionization from a diffuse radiation field far from a single star has not been examined. As shown above, accounting for stellar radiation feedback locally alone is insufficient to describe the long-term evolution of an isolated dwarf galaxy. To explore the cause of this difference we present two panel-plots in Figure~\ref{fig:panel1} and Figure~\ref{fig:panel2} which compare the gas number density (left), temperature (middle), and hydrogen ionization fraction (right) in edge-on slices in each of our simulations at two different times.\footnote{See ({\it insert link here once posted}) for a movie of this comparison.}

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{\textbf{In the final version of this figure I will add labels for each row marking the simulation, and a line giving the scale of each figure.But for now: TOP: Fiducial. MIDDLE: noRT. BOTTOM: shortrad} Edge-on slices of each simulation showing gas number density (left), temperature (middle), and hydrogen ionization fraction (right) 17~Myr after the formation of the first star in each run.}
\label{fig:panel1}
\end{figure*}

@@

\section{Conclusion}  \label{sec:conclusion}
In agreement with previous works we find that stellar radiation feedback is effective in regulating star formation and driving outflows in our simulations of an isolated, low mass dwarf galaxy. Simulations run without ionizing radiation feedback have star formation rates a factor of five higher than our fiducial simulation. In spite of this difference, SNe in our fiducial run are capable of driving larger galactic outflows, aided significantly by the ionizing radiation feedback.

We also demonstrate that simple prescriptions which treat stellar radiation feedback locally do not fully reproduce the same evolution as our fiducial model. Our simulation with radiation localized to 20~pc around each star particle does effectively regulate star formation on short time scales, predominately by quickly destroying cold, dense gas around young, hot stars. However, this model does not drive the significant outflows seen in our fiducial simulation. Long-range ionizing radiation is important for carving channels allowing the release of significant amounts mass and metals from SNe. Our simulation with localized radiation feedback retains metals at a significantly higher fraction than expected observationally for low mass dwarf galaxies.

Finally, we note that we have performed this experiment on only one possible type of galaxy. Its low virial temperature ($\sim10^{4}$~K) makes this galaxy particularly sensitive to the effects of stellar feedback and ionizing radiation in particular. Examining the role of long-range, diffuse stellar ionizing radiation on star formation and galactic winds in more massive galaxies is an important avenue of future research.
